---
title: "Project_Writeup"
author: "Tariq Alhindi"
date: "Sunday, May 24, 2015"
output: html_document
---

Loading required libararies
```{r}
library(caret)
library(randomForest)
```


### Preprocessing
Reading datasets and removing columns that have all missing values
```{r}
rawTrain <- read.csv("pml-training.csv")
training <- read.csv("pml-training-cleaned.csv")
#training <- Filter(function(x){!all(is.na(x))}, rawTrain)

rawTest <- read.csv("pml-testing.csv")
testing <- Filter(function(x){!all(is.na(x))}, rawTest)
```


Removing the first seven columns as they have irrelavant information to the model
e.g. username, timestamp...
```{r}
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]
```


### Data Split and Cross Validation
Spliting the training data into new Training and new Testing sets, where the new Training set will consisting of 75% of the original training data, which will be used to build the model. The remaining 25% will be used to test the accuracy of the model before applying it to the 20 test cases that we want to predict.
```{r}
inTrain     <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
newTraining <- training[inTrain, ] 
newTesting  <- training[-inTrain, ]
```


### Exploratory Plots

```{r}
plot(newTraining$classe, col="red", main="Histogram of the Classe variable in the new Training set", xlab="levels of classe variable", ylab="Frequency")
```

### Building the Model and validation:
We use the random forest approach to build the model that consist of 53 predictors. Random forrest was chosen because it had an accuracy > 99% and was eventually able to predict all 20 test cases succesfully upon submission. 
```{r}
modfit <- randomForest(classe ~. , data=newTraining, method="class")
predictions <- predict(modfit, newTesting, type = "class")
```

### Sample Error
As can be seen below, the error is lower than 1% which gives high credibility for our model
```{r}
confusionMatrix(predictions, newTesting$classe)
```

Applying the model to the 20 test cases that we want to predict
```{r}
TestingDatasetPrediction <- predict(modfit, testing, type = "class")
```

Below are the classes for the 20 test cases of desire
```{r}
as.character(TestingDatasetPrediction)
```


### Submission
In the end, we use the following function provided in the class to generate the 20 files needed to submit the programming part of the assignment
```{r}
answers = as.character(TestingDatasetPrediction)

pml_write_files = function(x){
     n = length(x)
     for(i in 1:n){
          filename = paste0("problem_id_",i,".txt")
          write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
     }
}

pml_write_files(answers)
```



